---
title: "Lecture12 - StructBioInf - Pt2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Structural Bioinformatics - Part 2

## Obtaining/preparing a protein sequence for docking analysis.
### 1. Use Bio3D to draw the pdb files for HIV protease HSD1.
```{r, include=FALSE}
library(bio3d)
```

```{r}
hivpdb <- get.pdb("1hsg")
# Next use the read.pdb()function to read this PDB file into R so we can prepare it for further analysis.
hiv <- read.pdb(hivpdb)
hiv
```
#### Question 1:
What is the name of the two non protein resid values in this structure? What does resid correspond to and how would you get a listing of all reside values in this structure? 

* Resid corresponds to the R groups on the moelcule
* The two non-protein residuals are H2O, and MK1.
* We can get a complete list of all the residuals by using the following command:

```{r}
View(hiv$atom$resid)
```


### 2. Prepare the file for analysis by trimming out specific things

```{r}
prot <- atom.select(hiv, "protein", value = TRUE)
prot
lig <- atom.select(hiv, "ligand", value = T)
lig
write.pdb(prot, file="1hsg_protein.pdb")
write.pdb(lig, file="1hsg_ligand.pdb")
```

### 3. Add charges, hydrogens, etc for ADT
(Done in ADT)

### 4. Use AutoDock Vina
```{r}
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")
```


### Compare structure
```{r}
res <- read.pdb("all.pdbqt", multi=TRUE) 
  # multi = T because this file has 14 different configuration models
ori <- read.pdb("ligand.pdbqt")
rmsd(ori, res)
```

## Normal Mode Analysis in Bio3D


THe above analyses predicts how a ligand fits into a protein but only describes them as static. In reality, molecules are constantly undergoing random motion. Analysis techniques that account for this, Normal mode analysis for example, allows us to predict the flexibility of proteins/molecules and model with that.

```{r}
pdb <- read.pdb("1hel")
modes <- nma( pdb )
m7 <- mktrj(modes, mode=7, file="mode_7.pdb")
plot(modes)
```

```{r}
library("bio3d.view")
view(m7, col=vec2color(rmsf(m7)))
```

