---
title: "BGGN213_HWK1"
author: "ACGeffre"
date: "4/26/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Lecture 6

## Section 1
```{r}
# (A. Can you improve this analysis code?
df1 <- data.frame(a=1:10, b=seq(200,400,length=10),c=11:20,d=NA)
df <- data.frame(a=1:10, b=seq(200,400,length=10),c=11:20,d=NA)

df$a <- (df$a - min(df$a)) / (max(df$a) - min(df$a))
df$b <- (df$b - min(df$a)) / (max(df$b) - min(df$b))
df$c <- (df$c - min(df$c)) / (max(df$c) - min(df$c))
df$d <- (df$d - min(df$d)) / (max(df$a) - min(df$d)) 

# Improvement
datafr <- function(x) {
  (x - min(x))/(max(x) - min(x))
}

df2 <-data.frame(a=1:10, b=seq(200,400,length=10),c=11:20,d=NA)
df2$a <- datafr(df2$a)
df2$b <- datafr(df2$b)
df2$c <- datafr(df2$c)
df2$d <- datafr(df2$d)
```

-----

## Section B
Use bio3d package to complete questions 1-6.
```{r}
#install.packages("bio3d")
library(bio3d)

# Can you improve this analysis code?
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b # variable atom has nested factors in it
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

# Question 1
Objects created by pdb are lists of class "pdb" and "sse" (sse appears to refer to inclusion of helix and sheet components.

# Question 2
trim.pdb() creates an abbreviated version of a pdb object - a subset of atoms according to the parameters loading through the built-in atom.select() function.

# Question 3
To turn off the rectangles from the plot, we use the "top" and "bot" input parameters:
[plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor", top = FALSE, bot = FALSE)]

These rectangles are derived from the "sse", or secondary structure information and describe the secondary structures present in the protein.
 
# Question 4
We could use a dendrogram to assess the relatedness of the structures of interactions sites across proteins.(We could also use the fancy overlay plots Dr. Grant showed us in lecture - I think it is struct.aln()?)

# Question 5
4AKE and 1E4Y are more similar in their b-factor trends, according to this dendrogram (built by hclust()), which plots each on according to their similarity (using dist()) in the b region (the part chunked out from each protein using atom$b and bound together in a df using rbind()). 
```{r}
hc <- hclust(dist(rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```


---------------
## Homework for 04/26/2019
# Question 6

"Write your own function starting from the code above that analyzes protein drug
interactions by reading in any protein PDB data and outputs a plot for the specified
protein."

# Baseline code
# Can you improve this analysis code?
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b # variable atom has nested factors in it
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")

```{r}
# Write a function that does what the above code did but in less lines and with more repeatability
protdrug <- function (x, y) {  
    # X = "protein name"", Y = "desired factor to measure""
  prot <- read.pdb(x)  
    # read in the protein file as an object
  prot_chain <- trim.pdb(prot, chain = "A", elety = "CA")  
    # Selects the A-chain, for what appears to be the element Calcium?
  protb <- prot_chain$atom$b
    # Chunks out the nested variable b from atom
  plotb3(protb, sse = prot_chain, typ = "l", ylab = y)
    # Plots this all with the y-axis label from input item y
}
```

```{r}
# Test it with a different input
test <- read.pdb("4q21") # kinase with drug
test.chainA <- trim.pdb(test, chain="A", elety="CA")
test.b <- test.chainA$atom$b
plotb3(test.b, sse=test.chainA, typ="l", ylab="Bfactor")
# Produces a test plot from 4q21
```

```{r}
protdrug("4q21", "Bfactor")  
# This appears to produce the same chart as the above code, so it works okay
```

```{r}
# test it with a different input, again
test2 <- read.pdb("4lhy") # kinase with drug
test2.chainA <- trim.pdb(test2, chain="A", elety="CA")
test2.b <- test2.chainA$atom$b
plotb3(test2.b, sse=test2.chainA, typ="l", ylab="Bfactor")
# Produces a test2 plot from 4lhy
```

```{r}
protdrug("4lhy", "Bfactor")  
# This appears to also produce the same chart as the above code.

```

```{r}
test3 <- read.pdb("1AKE") # kinase with drug
test3.chainA <- trim.pdb(test3, chain="A", elety="CA")
test3.b <- test3.chainA$atom$b
plotb3(test3.b, sse=test3.chainA, typ="l", ylab="Bfactor")
# Produces a test3 plot from 1AKE
```

```{r}
protdrug("1AKE", "Bfactor")  
# This appears to also, also produce the same chart as the above code.
```


