---
title: "Lecture 11 - Struvtural Bioinformatics"
author: "ACGeffre"
date: "5/8/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Structural Bioinformatics
This is the study of the structure of proteins, especially the volutionary lineages and functions of proteins that we can derive from their sturcture. 

## Why do we use structural informatics?
1. Sequence > Structure > Function
  + Structure determines function, so understanding structure helps our understanding of   function

2. Structure is more conserved than sequence
  + Structure allows identification of more distant evolutionary
relationships

3. Structure is encoded in sequence
 + Understanding the determinants of structure allows design and
manipulation of proteins for industrial and medical advantage.

## What are the goals fo sturct. bioinform.?
1. Visualization
  + Look at the actual structure
2. Analysis
  + 
3. Comparison
  + 
4. Prediction 
  + Can we predict function from the structure?
5. Design
  + Can we use what we know about port. structure to design new biomolecular motors/robots?

## Let's look at PDB experiments
We download the summary statistics of PDB data distribution by experiment- and molecule type. We download statistis data from PDB (05.08.2019).

```{r}
pdbstat <- read.csv("DataExportSummary.csv")
summary(pdbstat)

```


### What percentage of the structures use X-ray crystallography to solve?

```{r}

```


### What proportion of the structures are proteins?

```{r}

```

