---
title: "Lecture 11 - Struvtural Bioinformatics"
author: "ACGeffre"
date: "5/8/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Structural Bioinformatics
This is the study of the structure of proteins, especially the volutionary lineages and functions of proteins that we can derive from their sturcture. 

## Why do we use structural informatics?
1. Sequence > Structure > Function
  + Structure determines function, so understanding structure helps our understanding of   function

2. Structure is more conserved than sequence
  + Structure allows identification of more distant evolutionary
relationships

3. Structure is encoded in sequence
 + Understanding the determinants of structure allows design and
manipulation of proteins for industrial and medical advantage.

## What are the goals fo sturct. bioinform.?
1. Visualization
  + Look at the actual structure
2. Analysis
  + 
3. Comparison
  + 
4. Prediction 
  + Can we predict function from the structure?
5. Design
  + Can we use what we know about port. structure to design new biomolecular motors/robots?

## Let's look at PDB experiments
We download the summary statistics of PDB data distribution by experiment- and molecule type. We download statistis data from PDB (05.08.2019).

```{r}
pdbstat <- read.csv("DataExportSummary.csv")
summary(pdbstat)
str(pdbstat)
```

### Question 1
#### What percentage of the structures use X-ray crystallography to solve?

```{r}
percent <- ((pdbstat$Total)/sum(pdbstat$Total))*100 
  # For each row in $Total, make a percentage
percent 
namespdb <- pdbstat$Experimental.Method
  # Name vector from the experimental type
names(percent) <- namespdb
  # Feed names onto percent vector
percent
  # Ta-da!
```


#### What proportion of the structures are proteins?

```{r}

```


### Question 2

Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

* THere are 1157 protease structures currently in PDB

## Section 2 - Visualizing HIV-1 Protease
### Using Bio3D

```{r, include = FALSE}
#install.packages("bio3d")
library(bio3d)
```

```{r}
pdb <- read.pdb("1hsg")
# Grabbing the HSG1 pdb file from the internets
pdb
```

Attributes describes the data neted in the read-in pdb file.
```{r}
attributes(pdb)
```

aa321 allows simplification of the seqres datum
```{r}
aa321(pdb$seqres) # Prints AA sequence as single letters
```

```{r}
pdb$atom # atom is stored as a dataframe within a dataframe
```
```{r}
pdb$atom[1:2, c("eleno", "elety", "x","y","z")]
## eleno elety x y z
## 1 1 N 64.080 50.529 32.509
## 2 2 CA 64.044 51.615 33.423
# Note that individual $atom records can also be accessed like this
pdb$atom$elety[1:2]
```

#### Atom selection examples
```{r}
# Select residue 10
atom.select(pdb, resno = 10)

atom.select(pdb, resno = 10, value = TRUE)
```

### Select a subset of atoms and write a new pdb
```{r}
pdb1 <- atom.select(pdb, "protein")
write.pdb(pdb1, "protein-only_hsg1.pdb", pdb1$xyz)

liga <- atom.select(pdb, "ligand")
write.pdb(prot, "ligand-only_hsg1")
```

```{r}
#install.packages("rgl")
library(rgl)
```

```{r}
install.packages("devtools")
library(devtools)

devtools::install_bitbucket("Grantlab/bio3d-view")
install.packages("bio3d.view")
library(bio3d.view)

```


```{r}

```

