---
title: "BGGN213_Lecture9"
author: "ACGeffre"
date: "May 1st 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# PCA continued (in Lecture 9)
## Scaling - a PCA issue

We can use 'prcomp(x, scale=TRUE)' to scale data in PCAs. WHy scale our data?  

Variables may be in very different units (e.g. for the mtcars dataset, we have variables mpg (distance), and cylinders (absolute count) - these are both very different variables)


# Hands-on Lab excersize
## Reading in the data
We are using a data set of women from Wisconsin, who had breat biopsies to assay for cancer.Thsi dataset includes the patient ID, diagnosis (b or m) and 31 continuous variables describing physical properties fo the biopsy.  

We will call this dataset will be called "wisc"

```{r}
wisc <- read.csv("WisconsinCancer.csv")
head(wisc) # Describes the data
length(wisc$id)  # how many patients?
table(wisc$diagnosis) # How many benign (b) and malignant (m) disgnoses?
length(grep("_mean", colnames(wisc))) # how many columns have the phrase "_mean" in them?
```

## Prepare the data for PCA 
To ready out dataframe for PCA, we need to convert it to a matrix, and also make sure only numeric data exists therein. (e.g. pop off non-numeric columns, like ID and diagnosis, and the empty x column full of NAs at the end).

```{r}
# make a matrix from all the rows, but only columns 3-32 of the datafram wisc
wiscdat <- as.matrix(wisc[,3:32]) 
```

Now we will import the patient IDs onto the rownames of the matrix. This prevents the need to have columns with strings in them.

```{r}
row.names(wiscdat) <- wisc$id # Pop the names from original set to the new matrix
head(wiscdat)  # Now the matrix has row names that correspond to the patient IDs
```

For visualizing cluster patterns later, we are going to store diagnosis as a vecotr to be overlayed onto our PCA plots later.

```{r}
diagnosis <- wisc$diagnosis
# This stores the diagnosis for each patient in a vector we can use with PCA later
```

## Run the PCA

Firstly, we discussed that data variables with very different units/variances can be difficult to collapse in PCA.  Let's check our data for this issue:

### Check the data
```{r}
colMeans(wiscdat) # find the means for each column
hist(colMeans(wiscdat), breaks = 20)
```

There are some means that are very different from the others - suggests we might want to scale.


```{r}
wiscsd <- apply(wiscdat,2,sd) # apply the sd() function over all columns
hist(wiscsd, breaks = 20)
```

Ditto the std. deviation of the data - also suggests we might want to scale.

We can scale before hand, using scale(), or feed this argument into prcomp (see below).

### PCA the data
```{r}
wiscpr <- prcomp(wiscdat, scale. = TRUE) # scale them shits
wiscpr
summary(wiscpr)  # summarize them shits
```

## Plot the data

### Biplot
Very information intense! Hard to interpret!
```{r}
biplot(wiscpr)
````

### Base R
This is similar to what we used before.
```{r}
plot(wiscpr$x[,1], wiscpr$x[,2], col = diagnosis, pch=16,
     xlab=paste0("PC1 (", ((wiscpr$sdev^2)*100)[1]/sum((wiscpr$sdev)^2), "%)"),
     ylab=paste0("PC2 (",((wiscpr$sdev^2)*100)[2]/sum((wiscpr$sdev)^2), "%)"))

# Note that we use the "diagnosis" vector we created earlier
# Also note that we pull out the varience explained by the following:
# ((wiscpr$sdev^2)*100)[1]/sum((wiscpr$sdev)^2, where the [1] indicates teh first PC
```

What about the other PCs?  PC3 explained similar varience as PC2.
```{r}
plot(wiscpr$x[,1], wiscpr$x[,3], col = diagnosis, pch=16,
     xlab=paste0("PC1 (", ((wiscpr$sdev^2)*100)[1]/sum((wiscpr$sdev)^2), "%)"),
     ylab=paste0("PC2 (",((wiscpr$sdev^2)*100)[3]/sum((wiscpr$sdev)^2), "%)"))

```

These look a bit same-y, but the cut between the two groups isn;t as clean - it's better to use the PC1 and PC2 model, to predict if new data would indicate benign or malignant tissue biopsies.

## Thinking about Varience (Scree Plots)

Let's look at how much varience each of our PCs explain graphically (with scree plots!)

```{r}
# Makes a varience vector
prvar <- ((wiscpr$sdev^2)*100)
```

```{r}
# Makes a percent varience explained vector
pve <- (prvar/sum((wiscpr$sdev)^2))
```


### Base R
Let's make one with our big garrish calc. 
```{r}
hist(((wiscpr$sdev^2)*100)/sum((wiscpr$sdev)^2), breaks = 100)
```

### Also Base R
Now let's try with our pretty vectors.  
```{r}
plot(pve, xlab = "Principal Component", # Uses our PVE vector generated above
     ylab = "Proportion of Variance Explained (%)", 
     ylim = c(0, 100), type = "o") # Makes a dot and line chart with type = "o"
```

### More streamlined base R
```{r}
# Alternative scree plot of the same data, note data driven y-axis
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```

### GGplot using factoextra!

```{r}
# ggplot based graph
#install.packages("factoextra") # comment out install.packages() after we install them
library(factoextra)
fviz_eig(wiscpr, addlabels = TRUE)
```

Slick, AF, my friends - slick AF. factoextra is kind of a beast when it comes to scree plots.

## Check out the laoding factors of PCs

We can tell the contribution of a given variable by determining how much it contributes to the PC - we do this by check the loading factor.
```{r}
# rotatiom is the loading factor, 1 is the PC id
wiscpr$rotation["radius_mean",1] 
wiscpr$rotation["smoothness_se",1]
```

How can we look for the loading factor that contributes the MOST to PC1? Recall larger loading factor means a greater contribution!

For example, this would translate to:ß
PC1 score = (variable 1 value X loading factor of variable 1(PC1)) + 
            (variable 2 value X loading factor of variable 2(PC1)) + 
            ... + 
            (variable n value * loading factor of variable n(PC1))

```{r}
sort(   # sorts the data
  abs(  # give the absolute value 
    wiscpr$rotation[,1]))  # for the first PC in wiscpr

```

# Hierarchical Clustering
## Data preparation
As we did with the prcomp function, we need to construct a scaled data set.  We can use scale(), which scales and centers matrix [and matrix-like] objects.
```{r}
datascaled <- scale(wiscdat) # scale our matrix data
```

Next, we will calculate the Euclidean distance between our data poits
```{r}
distdata <- dist(datascaled, method = "euclidean") 
```

Aside: Let's try again with factorextra's dist() function
```{r}
fviz_dist(distdata, order = TRUE, show_labels = TRUE, lab_size = NULL,
  gradient = list(low = "red", mid = "white", high = "blue"))

#oly Geeze this is computationally itensive, comparatively!

```

### Do cluster analysis using "complete" linkage
```{r}
wisc.hclust <- hclust(distdata, method = "complete")
```

```{r}
plot(wisc.hclust)
abline(h=19, col = "red")
```

CLuster our data into four groups using cutree:
```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k = 4)
table(wisc.hclust.clusters, diagnosis)
```

## Combine clustering with PCA

```{r}
wisc.pr.hclust <- hclust(dist(wiscpr$x[,1:7]), method = "ward.D2")  
# ook more into the methods - I;m not sure why to use Ward D2 here.
# Here, we'll use PCs 1 through 7 because they explain ~ 90% of varience; this isn't always the right option, though.
plot(wisc.pr.hclust)
```

```{r}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
table(grps, diagnosis)
```

```{r}
plot(wiscpr$x[,1:2], col=grps)
```

THe col= input changes the ordering of the coloration, but the groups derived from the PCA seem to distinguish the B and M patients well (see below as reference). 

```{r}
plot(wiscpr$x[,1:2], col=diagnosis)
```

### Pretty 3D PCA plot for fun
Here is the code, although I can't load it because my R version is too new.
```{r}
#install.packages("rgl")
#install.packages("XQuartz")
#library(rgl)
#plot3d(wisc.pr$x[,1:3], xlab="PC 1", ylab="PC 2", zlab="PC 3", cex=1.5, size=1, type="s", col=grps)
```

## Predicting diagnosis with PCA

We can use our PCA model to predict the diagnosis of new samples. Let's try with new sample input:
```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wiscpr, newdata=new)
npc
```

```{r}
plot(wiscpr$x[,1:2], col=diagnosis)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3) # Plot our two new patients 
text(npc[,1], npc[,2], c(1,2), col="white")
```
From this, we can see that patient 2 appears to have a biopsy indicating malignancy!  She should be considered for treatment!

Thanks PCA!!

# Session Deets
```{r}
sessionInfo()
```

