---
title: "BGGN213, S2019: Functions Homework"
author: "Amy Geffre"
date: "4/26/2019"
output:
  github_document:
    toc: true
    highlight: tango
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
#install.packages("bio3d")
library(bio3d)
```


# Homework for 04/26/2019

"Write your own function starting from the code above that analyzes protein drug
interactions by reading in any protein PDB data and outputs a plot for the specified
protein."

### Baseline code
**Can you improve this analysis code?**

```{r, message=FALSE}
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
```

```{r, fig.cap = "Wordy code output plots", fig.align="center"}
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b # variable atom has nested factors in it
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

par(mfrow=c(1,3))
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```


## **Let's write a simpler function!**
```{r}
# Write a function that does what the above code did but in less lines and with more repeatability
protdrug <- function (x, y) {  
    # X = "protein name"", Y = "desired factor to measure""
  prot <- read.pdb(x)  
    # read in the protein file as an object
  prot_chain <- trim.pdb(prot, chain = "A", elety = "CA")  
    # Selects the A-chain, for what appears to be the element Calcium?
  protb <- prot_chain$atom$b
    # Chunks out the nested variable b from atom
  plotb3(protb, sse = prot_chain, typ = "l", ylab = y)
    # Plots this all with the y-axis label from input item y
}
```


--------

## **Test the function** 
### *Part 1* - Make plot with normal wordy code
```{r,warning = FALSE, fig.cap = "The original code output", fig.align="center"}
# Test it with a different input
test <- read.pdb("4q21") # kinase with drug
test.chainA <- trim.pdb(test, chain="A", elety="CA")
test.b <- test.chainA$atom$b
plotb3(test.b, sse=test.chainA, typ="l", ylab="Bfactor")
```


### *Part 2* - See if my function output matches the wordy code
```{r, fig.cap = "My funtcion output", fig.align="center"}
protdrug("4q21", "Bfactor")  
```
It matches!


---------------


## **Check the Function**
### *Try 2*
```{r,warning = FALSE, fig.cap = "Original Code Output", fig.align="center"}
# test it with a different input, again
test2 <- read.pdb("4lhy") # kinase with drug
test2.chainA <- trim.pdb(test2, chain="A", elety="CA")
test2.b <- test2.chainA$atom$b
plotb3(test2.b, sse=test2.chainA, typ="l", ylab="Bfactor")
```


```{r, fig.cap = "My function output", fig.align="center"}
protdrug("4lhy", "Bfactor")  
```
It matches again


### *Try 3*
```{r, warning = FALSE, fig.cap = "Original code output", fig.align="center"}
test3 <- read.pdb("1AKE") # kinase with drug
test3.chainA <- trim.pdb(test3, chain="A", elety="CA")
test3.b <- test3.chainA$atom$b
plotb3(test3.b, sse=test3.chainA, typ="l", ylab="Bfactor")
```


```{r, fig.cap = "My function output", fig.align="center"}
protdrug("1AKE", "Bfactor")  
```
It matches again, again. 


--------


## *Final comment/question*
These results don't say I can be sure it would work on all pdb files.  How would we test stuff like that?

