---
title: "Lect7Stuff"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## More on Function writing

Revist function fro previous class (which I missed ;_;)

```{r}
source("http://tinyurl.com/rescale-R") # loads a suite of functions associated with "Rescale", from the internets
```

Let's test the **rescale()** function

```{r}
rescale(1:10) # relevels the numebers from 1 to 10 to fit betweeb 0 and 1. 
```

When writing functions we can include warning messages and/or stop the function from working in the instance of weird input.

E.G. we can use warning(), to give a warning but continue the function, or the stop() function to cease the function before finishing. 

**For example: baseline rescale()**
function(x, na.rm=TRUE, plot=FALSE, ...) {
  # Our rescale function from the end of lecture 9

  if(na.rm) {
    rng <-range(x, na.rm=TRUE)
  } else {
    rng <-range(x)
  }

  answer <- (x - rng[1]) / (rng[2] - rng[1])
  if(plot) { 
    plot(answer, ...) 
  }

  return(answer)
}

**Rescale with warnings:**
function(x, na.rm=TRUE, plot=FALSE, ...) {
  # Our rescale function from lecture 10

  if( !is.numeric(x) ) {
    stop("Input x should be numeric", call.=FALSE)
  }
  
  rng <-range(x, na.rm=TRUE)

  answer <- (x - rng[1]) / (rng[2] - rng[1])
  if(plot) { 
    plot(answer, ...) 
  }

  return(answer)
}
```{r}
# Compare outputs for weird inputs with both functions
rescale(c(1:10, "string"))

```
```{r}
rescale2(c(1:10, "string"))
# This error message is plainer english - perhaps easier to read.
```

Good function writing
• Understandable (remember that functions are for humans
and computers)
• Correct + Understandable = Obviously correct
• Use sensible names throughout. What does this code do?
• Good names make code understandable with minimal
context. You should strive for self-explanatory names

```{r}
# Write a function for us
# Create dataset for testing
x <- c( 1, 2, NA, 3, NA) 
y <- c(NA, 3, NA, 3, 4)

```

```{r}
is.na(x) & is.na(y) # looks for when both data sets are cuncidentally
sum(is.na(x) & is.na(y) )
# These results indicate that there is only onespot for in th enew DNA where thei;r seuquences are the same.
```


```{r}

both_na <- function(x,y) {
  sum(is.na(x) & is.na(y))
}
```

```{r}
both_na(c(NA, NA, NA), c(NA, NA, 1))
# the end  ouput is a rsult which desxfibes the two datas at which your views on both platforms are the same. 
```

```{r}
both_na2 <- function(x, y) {
 if(length(x) != length(y)) {
 stop("Input x and y should be the same length")
 }
 sum( is.na(x) & is.na(y) )
}
```



```{r}
both_na3 <- function(x, y) {
 if(length(x) != length(y)) {
 stop("Input x and y should be vectors of the same length")
 }

 na.in.both <- ( is.na(x) & is.na(y) )
 na.number <- sum(na.in.both)
 na.which <- which(na.in.both)  # where "which: finds positions that are true to this statement.
 message("Found ", na.number, " NA's at position(s):",
 paste(na.which, collapse=", ") )

 return( list(number=na.number, which=na.which) )
}
```


Write a grading function (as in lab handout for this week)
```{r}
# We need to select remove the lowest scores from each student's assignments
# After that we can calculate a mean score for each student and determine if they have passed. 

# Create "Student 1" and "Student 2" vectors to test our function with
st1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
st2 <- c(100, NA, 90, 90, 90, 90, 97, 80)

grade <- function(x) {
  (sum(x, na.rm = T) - min(x, na.rm = T))/(length(x)-1)
  }

grade(st1)
grade(st2)
```

Let's apply this to a dataframe.
```{r}
hwk <- read.csv("student_homework.csv", row.names = 1)

grade(hwk[1,])
apply(hwk, 1, grade) # use margin = 1 to apply across rows!
# apply(df, MARGIN = "", FUNCTION) is the basic format for apply

```
```{r}
# Sort by top performers in class
ans <- apply(hwk, 1, grade)
sort(ans, decreasing = T ) # sorts our vector by value in decreasing order
```

## More examples of functions

```{r}
source("https://tinyurl.com/rescale-R")
df1 <- data.frame(IDs=c("gene1", "gene2", "gene3"),
    exp=c(2,1,1),
    stringsAsFactors=FALSE)

df2 <- data.frame(IDs=c("gene2", "gene4", "gene3", "gene5"),
    exp=c(-2, NA, 1, 2),
    stringsAsFactors=FALSE)

# Make single vectors
x <- df1$IDs
y <- df2$IDs

intersect(x,y) # find gene ids that match between data sets

x %in% y # Finds all items in x, that are contained in y
x[x %in% y] # Lists the overlapping IDs

y %in% x # Finds all items in y, that are contained in y
y[y %in% x]

cbind(x[x %in% y], y[y %in% x])
# Make that shit a function
gene_intersect <- function(x, y) {
 cbind( x[ x %in% y ], y[ y %in% x ] )
}

gene_intersect(x,y)
# Makes a cute little table for this.
```

```{r}
# PRevious function doesn't work with dataframes, though.
gene_intersect2 <- function(df1, df2) {
 cbind( df1[ df1$IDs %in% df2$IDs, ], #assumes my df have an "ID" column
 df2[ df2$IDs %in% df1$IDs, "exp"] )
}
gene_intersect2(df1, df2)
# 

```
```{r}
gene_intersect3 <- function(df1, df2, gene.colname="IDs") {  
  # allows user to specify ID names through gene.colname input
 cbind( df1[ df1[,gene.colname] %in%
 df2[,gene.colname], ],
 exp2=df2[ df2[,gene.colname] %in%
 df1[,gene.colname], "exp"] )
}


```

